Supplementary material

### 2025-08-08 resubmission/04-analisis.qmd
Up carpeta resubmit

p6 <- ggpredict(
  m9.1,
  terms = c("merit_effort_mean", 
            sprintf("perc_inequality_mean [%.2f, %.2f, %.2f]", min_val, mean_val, max_val))
) %>% 
  plot() +
  scale_x_continuous(n.breaks = 5) +
  scale_color_manual(
    name = "Perception of inequality (BE)",
    values = c("#0571B0", "black", "#B2182B"),
    labels = c("Min", "Mean", "Max")
  ) +
  scale_fill_manual(
    name = "Perception of inequality (BE)",
    values = c("#0571B0", "black", "#B2182B"),
    labels = c("Min", "Mean", "Max")) +
  labs(
    x = "Merit: Effort (BE)",
    y = "Market justice preferences",
    title = "(A) People are rewarded by their efforts") +
  theme_ggdist()

p7 <- ggpredict(
  m9.2,
  terms = c("merit_talent_mean", 
            sprintf("perc_inequality_mean [%.2f, %.2f, %.2f]", min_val, mean_val, max_val))
) %>% 
  plot() +
  scale_x_continuous(n.breaks = 5) +
  scale_color_manual(
    name = "Perception of inequality (BE)",
    values = c("#0571B0", "black", "#B2182B"),
    labels = c("Min", "Mean", "Max")
  ) +
  scale_fill_manual(
    name = "Perception of inequality (BE)",
    values = c("#0571B0", "black", "#B2182B"),
    labels = c("Min", "Mean", "Max")) +
  labs(
    x = "Merit: Talent (BE)",
    y = "Market justice preferences",
    title = "(B) People are rewarded for their intelligence",
  caption = "Source: own elaboration with data from ELSOC 2016-2023 (n = 8460). Based in Model 10 and Model 11 in Table 4") +
  theme_ggdist()

@@

```{r echo=FALSE, results='asis', eval=FALSE, include=FALSE}
#| label: tbl-interactions2
#| tbl-cap: "Interactions for meritocracy, perceived economic inequality and market justice preferences"
#| tbl-cap-location: top

ccoef <- list(
  "(Intercept)" = "Intercept",
  "ola_num:perc_inequality_cwc" = "Wave x Perception inequality (WE)",
  "ola_num:merit_effort_cwc" = "Wave x Merit: effort (WE)",
  "ola_num:merit_talent_cwc" = "Wave x Merit: talent (WE)",
  "ola_num:perc_inequality_mean" = "Wave x Perception inequality (BE)",
  "ola_num:merit_effort_mean" = "Wave x Merit: effort (BE)",
  "ola_num:merit_talent_mean" = "Wave x Merit: talent (BE)"
    )

texreg::texreg(list(m10.1, m11.1, m12.1, m10.2, m11.2, m12.2),
               custom.model.names = c("Model 12", "Model 13", "Model 14", "Model 15", "Model 16", "Model 17"),
               caption = NULL,
               stars = c(0.05, 0.01, 0.001),
               custom.coef.map = ccoef,
               digits = 3,
               caption.above = T,
               custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. %stars. CWC = centered within group.",
               leading.zero = T,
               use.packages = F,
               booktabs = F,
               scalebox = 0.7,
               include.loglik = FALSE,
               include.aic = FALSE,
               center = T,
               custom.gof.names = c("BIC", 
                                    "Numb. obs.", 
                                    "Num. groups: individuals", 
                                    "Var: individuals (Intercept)",

                                    "Var: individuals, perception inequality cwc",
                                    "Var: individuals, wave",
                                    "Cov: individuals (Intercept), perception inequality cwc",
                                    "Cov: individuals (Intercept), wave",
                                    "Cov: individuals, perception inequality cwc, wave",
                                    "Var: Residuals",
                                    "Var: individuals, merit effort cwc", 
                                    "Cov: individuals (Intercept), merit effort cwc",
                                    "Cov: individuals, merit effort cwc, wave",

                                    "Var: individuals, merit talent cwc", 
                                    "Cov: individuals (Intercept), merit talent cwc",
                                    "Cov: individuals, merit talent cwc, wave",

                                    "Var: individuals, perception inequality mean", 
                                    "Cov: individuals (Intercept), perception inequality mean",
                                    "Cov: individuals, perception inequality mean, wave",

                                     "Var: individuals, merit effort mean", 
                                    "Cov: individuals (Intercept), merit effort mean",
                                    "Cov: individuals, merit effort mean, wave",

                                    "Var: individuals, merit talent mean", 
                                    "Cov: individuals (Intercept), merit talent mean",
                                    "Cov: individuals, merit talent mean, wave"
                                    ),
               custom.gof.rows = list("Controls"=c(rep("Yes",6))))

```{r}
#| label: fig-growth
#| out-width: '100%'
#| fig-asp: 1
#| fig-cap: "Time varying within effects of meritocratic perceptions on market justice preferences"
#| fig-cap-location: top
#| echo: false
#| results: asis
#| include: false
#| eval: false

p4 <- ggpredict(m11.2, terms = c("ola_num [all]", "merit_effort_mean")) %>% 
  plot() + 
  scale_x_continuous(n.breaks = 6) +
  scale_color_manual(
    name = "Merit:\nEffort",
    values = c("#D55E00", "#0072B2", "#009E73"),
    labels = c("-1 SD", "Mean", "+1 SD")
  ) +
  scale_fill_manual(
    name = "Merit:\nEffort",
    values = c("#D55E00", "#0072B2", "#009E73"),
    labels = c("-1 SD", "Mean", "+1 SD")) +
  labs(x = "Wave",
       title = "(A) People are rewarded for their efforts",
       color = NULL) +
  ggdist::theme_ggdist() +
  theme(legend.position = "top")

p5 <- ggpredict(m12.2, terms = c("ola_num [all]", "merit_talent_mean")) %>% 
  plot() +
  scale_x_continuous(n.breaks = 6) +
  scale_color_manual(
    name = "Merit:\nTalent",
    values = c("#D55E00", "#0072B2", "#009E73"),
    labels = c("-1 SD", "Mean", "+1 SD")
  ) +
  scale_fill_manual(
    name = "Merit:\nTalent",
    values = c("#D55E00", "#0072B2", "#009E73"),
    labels = c("-1 SD", "Mean", "+1 SD")
  ) +
  labs(
    x = "Wave",
    title = "(B) People are rewarded for their intelligence",
    caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 8460)"
  ) +
  ggdist::theme_ggdist() +
  theme(legend.position = "top")

ggpubr::ggarrange(p4, p5, common.legend=FALSE, ncol=1)

>>>>

p6 <- ggpredict(
  m9.1,
  terms = c("merit_effort_mean", 
            sprintf("perc_inequality_mean [%.2f, %.2f]", min_val, max_val))
) %>% 
  plot() +
  scale_x_continuous(n.breaks = 5) +
  scale_color_manual(
    name = "Perception of inequality (BE)",
    values = c("#0571B0",  "#B2182B"),
    labels = c("Min",  "Max")
  ) +
  scale_fill_manual(
    name = "Perception of inequality (BE)",
    values = c("#0571B0",  "#B2182B"),
    labels = c("Min",  "Max")) +
  labs(
    x = "Merit: Effort (BE)",
    y = "Market justice preferences",
    title = "(A) People are rewarded by their efforts") +
  theme_ggdist()

p7 <- ggpredict(
  m9.2,
  terms = c("merit_talent_mean", 
            sprintf("perc_inequality_mean [%.2f, %.2f]", min_val, max_val))
) %>% 
  plot() +
  scale_x_continuous(n.breaks = 5) +
  scale_color_manual(
    name = "Perception of inequality (BE)",
    values = c("#0571B0",  "#B2182B"),
    labels = c("Min",  "Max")
  ) +
  scale_fill_manual(
    name = "Perception of inequality (BE)",
    values = c("#0571B0",  "#B2182B"),
    labels = c("Min",  "Max")) +
  labs(
    x = "Merit: Talent (BE)",
    y = "Market justice preferences",
    title = "(B) People are rewarded for their intelligence",
  caption = "Source: own elaboration with data from ELSOC 2016-2023 (n = 8460). Based in Model 10 and Model 11 in Table 4") +
  theme_ggdist()

@@

### 2025-08-08 resubmission/05-discussion.qmd
Up carpeta resubmit

The first set of hypotheses suggested that perceptions of economic inequality were strongly linked to preferences for market justice. The findings demonstrated that individuals perceiving higher income disparities showed less support for the idea that the access to core social services depends on individual income, aligning with theories emphasizing that awareness of inequality could fuel a critical view of such policies [@Castillo2012a_justice; @mijs_paradox_2021; @garcia-sanchez_vicious_2019].  In particular, the negative within-individual effect observed over time ($H1_{b}$) suggests that as inequality becomes more salient for some individuals, their support for market justice declines, potentially due to rising distrust in market mechanisms or growing disillusionment with the fairness of the system. However, although the effect of perceived inequality at the between-individual level is also negative, it is not statistically significant. Therefore, we cannot conclude that individuals who, on average, perceive higher inequality are systematically less supportive of market justice. As a result, hypothesis $H1_{a}$ does not receive empirical support.

>>>>

The first set of hypotheses proposed that perceptions of economic inequality are strongly associated with preferences for market justice. In the between-individual specification, the effect of perceived inequality is not statistically significant. Therefore, we cannot conclude that individuals who, on average, perceive higher levels of inequality are systematically less supportive of market justice. As a result, hypothesis $H1_{a}$ is not supported by the data. However, in the within-individual specification, the findings reveal a significant negative association: individuals who, at a given point in time, perceive higher income disparities are less supportive of the idea that access to core social services should depend on individual income. This result aligns with theories suggesting that increased awareness of inequality fosters a more critical stance toward market-based distributive arrangements [@Castillo2012a_justice; @mijs_paradox_2021; @garcia-sanchez_vicious_2019]. Such a pattern reflects a moral economy logic in which perceptions of systemic unfairness undermine the legitimacy of existing distributions and strengthen demands for greater equity. Specifically, the negative within-individual effect over time ($H1_{b}$) suggests that as inequality becomes more salient for some individuals, their support for market justice declines—possibly due to growing distrust in market mechanisms or increasing disillusionment with the perceived fairness of the system.

### 2025-08-08 resubmission/07-supplementary-material.qmd
Up carpeta resubmit

>>>>

```{r echo=FALSE, results='asis'}
#| label: tbl-interactions2
#| tbl-cap: "Growth curves for meritocracy, perceived economic inequality and market justice preferences"
#| tbl-cap-location: top

ccoef <- list(
  "(Intercept)" = "Intercept",
  "ola_num:perc_inequality_cwc" = "Wave x Perception inequality (WE)",
  "ola_num:merit_effort_cwc" = "Wave x Merit: effort (WE)",
  "ola_num:merit_talent_cwc" = "Wave x Merit: talent (WE)",
  "ola_num:perc_inequality_mean" = "Wave x Perception inequality (BE)",
  "ola_num:merit_effort_mean" = "Wave x Merit: effort (BE)",
  "ola_num:merit_talent_mean" = "Wave x Merit: talent (BE)"
    )

texreg::texreg(list(m10.1, m11.1, m12.1, m10.2, m11.2, m12.2),
               custom.model.names = c("Model 12", "Model 13", "Model 14", "Model 15", "Model 16", "Model 17"),
               caption = NULL,
               stars = c(0.05, 0.01, 0.001),
               custom.coef.map = ccoef,
               digits = 3,
               caption.above = T,
               custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. %stars. CWC = centered within group.",
               leading.zero = T,
               use.packages = F,
               booktabs = F,
               scalebox = 0.7,
               include.loglik = FALSE,
               include.aic = FALSE,
               center = T,
               custom.gof.names = c("BIC", 
                                    "Numb. obs.", 
                                    "Num. groups: individuals", 
                                    "Var: individuals (Intercept)",

                                    "Var: individuals, perception inequality cwc",
                                    "Var: individuals, wave",
                                    "Cov: individuals (Intercept), perception inequality cwc",
                                    "Cov: individuals (Intercept), wave",
                                    "Cov: individuals, perception inequality cwc, wave",
                                    "Var: Residuals",
                                    "Var: individuals, merit effort cwc", 
                                    "Cov: individuals (Intercept), merit effort cwc",
                                    "Cov: individuals, merit effort cwc, wave",

                                    "Var: individuals, merit talent cwc", 
                                    "Cov: individuals (Intercept), merit talent cwc",
                                    "Cov: individuals, merit talent cwc, wave",

                                    "Var: individuals, perception inequality mean", 
                                    "Cov: individuals (Intercept), perception inequality mean",
                                    "Cov: individuals, perception inequality mean, wave",

                                     "Var: individuals, merit effort mean", 
                                    "Cov: individuals (Intercept), merit effort mean",
                                    "Cov: individuals, merit effort mean, wave",

                                    "Var: individuals, merit talent mean", 
                                    "Cov: individuals (Intercept), merit talent mean",
                                    "Cov: individuals, merit talent mean, wave"
                                    ),
               custom.gof.rows = list("Controls"=c(rep("Yes",6))))
